```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "hw2"
author: "Haley Humphries haley22"
date: "02/07/20"
---

This homework is about syntax and basic plotting functions in R, and is **due by the end of Feb 7**. 

You can use the RMarkDown file with these homework instructions as a template for your work. 
Make sure to replace "Your Name" and "your_github_username" in the YAML with your name and Github username.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.










**Example answer**:
In the RMarkDown file, type in
<pre><code>```{r}
a0 <- 1
a0
```</code></pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 24 and a low of 15 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
temp = c(24, 15)
```
```{r}
cel = 5/9 * (temp-32)
cel
```

2. Create a vector of length 200. The odd entries should be 3 and even entries should be 1. Storage the created in a variable named `vec200` and print.
```{r}
vec200 <- 0
for (i in 1:200)
{
  if((i%%2)==0)
    vec200[i] <- 1
  else
    vec200[i] <- 3
  print(vec200[i])
}
```

3. How many perfect squares are there between 1 and 2001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
numPerfect <- 0
for(i in 1:2001)
{
  J <- 1
  while((J*J)<=i)
  {
    if((J*J)==i)
      numPerfect=numPerfect+1
    J <- J+1
  }
  i<- i+1
}
print(numPerfect)
```

## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by first installing (using `install.packages('ggplot2')`) and then loading the package `ggplot2` (using `library(ggplot2)`) into your R session. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.

```{r}
library(ggplot2)
```

Most exercises are from the ggplot2 book section 2.2.1. You might find it helpful for working on the homework to read through this section. *However, all questions can be solved by using just the basic R graphics functions.*

1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.
```{r}
top3 <- head(mpg[order(mpg$hwy, decreasing=T),], n=3)
top3
```

2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.
```{r}
str(mpg)
```
```{r}
numCompact <- subset(mpg, class == "compact")
```
```{r}
summary(numCompact)
```
Length is 47.

3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.
```{r}
x <- mpg$hwy
y <- mpg$cty
```

```{r}
ggplot(mpg, aes(x=hwy, y=cty)) + geom_point(shape=16, size=3) + theme_bw()
```
Highway mileage is better than city mileage.

4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. Explain your findings.

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.

```{r}
cars2008 <- subset(mpg, year== 2008)
cars1999 <- subset(mpg, year== 1999)
ggplot(cars2008, aes(x = cty)) + 
  geom_histogram(binwidth = 1) +
  ggtitle("binwidth = 1")
  ggplot(cars1999, aes(x = cty)) + 
  geom_histogram(binwidth = 1) +
  ggtitle("binwidth = 1")
  ggplot(cars2008, aes(x = hwy)) + 
  geom_histogram(binwidth = 1) +
  ggtitle("binwidth = 1")
  ggplot(cars1999, aes(x = hwy)) + 
  geom_histogram(binwidth = 1) +
  ggtitle("binwidth = 1")
```

```{r}
mean(cars2008$cty)
mean(cars2008$hwy)
mean(cars1999$cty)
mean(cars1999$hwy)
```

By creating these graphs and looking at the means you can see that nothing really improved over those 9 years and city mileage even decreased.
